\textbf{\supplement}\vskip 2ex \mytitle\vskip 2ex

### 2023-01-31 response-to-referees/paper-seb-supplementary-material (original).tex
More

\iftrue
    \PassOptionsToPackage{hyphens}{url}
    \documentclass[10pt,a4paper]{article}
    \usepackage[text={6.5in,10in},centering]{geometry}    
    \usepackage{url}
    \usepackage{doi}
    \usepackage{longtable}
    \usepackage{color}
    \usepackage{amsmath,amssymb}
    \usepackage{fancyhdr}
    \usepackage{graphicx}
\usepackage{amssymb}
\usepackage{epstopdf}
    \def\httpURL#1{\href{http://#1}{\textcolor{blue}{#1}}}
    \def\LONGhttpURL#1#2{\href{http://#1}{\textcolor{blue}{#2}}}    
    \input paper-seb-macros.tex
\fi

@@

\input generated/page-lengths.tex

@@

% pageref gives us a hyperlink which looks unsightly, so we do it this way...
\makeatletter
\def\rememberpage{\immediate\write\@auxout{\bslash gdef\bslash SupplementalLastPage{\thepage}}}
\makeatother

@@

% this will be used to save cross references that can be used in the main paper too
\newwrite\infofileForMain
\immediate\openout\infofileForMain=generated/info-for-main.tex

%\makeatletter
%% saves name-value pairs in both the supplementary material .aux file and the info for the main file
%\def\remembernamevalue#1#2{%
%	\immediate\write\@auxout{\@backslashchar gdef\@backslashchar #1{#2}}%
%	\immediate\write\infofileForMain{\@backslashchar gdef\@backslashchar #1{#2}}%
%}
%\makeatother

\input generated/info-for-supplement.tex
\newwrite\infofileForSupplement
\immediate\openout\infofileForSupplement=generated/info-for-supplement.tex

@@

\makeatletter
\def\globalLabel#1{% we also need journal code-policies to be used in the main paper...
\typeout{Global globalLabel: #1}%
\immediate\write\infofileForMain{\bslash newlabel{#1}{{\arabic{section}}{\thepage}{DUMMYA}{DUMMYB}{DUMMYC}}}
% as well as in the current document....
\immediate\write\@auxout{\bslash newlabel{#1}{{\bslash arabic{section}\dotalph{subsection}}{\thepage}{AAA}{BBB}{CCC}}}
}
\def\globalTableLabel#1{% we also need journal code-policies to be used in the main paper...
\typeout{Global globalTableLabel: #1}%
\immediate\write\infofileForMain{\bslash newlabel{#1}{{\arabic{table}}{\thepage}{DUMMYA}{DUMMYB}{DUMMYC}}}
% as well as in the current document....
\immediate\write\@auxout{\bslash newlabel{#1}{{\bslash arabic{table}\dotalph{subsection}}{\thepage}{AAA}{BBB}{CCC}}}
}
\def\globalLabelValue#1#2{% we also need journal code-policies to be used in the main paper...
\typeout{Global globalLabelValue: #1 := #2}%
\immediate\write\infofileForMain{\bslash newlabel{#1}{{#2}{\thepage}{AAA}{BBB}{CCC}}}
}
\makeatother

@@

\noindent\vbox{% The table of contents
\def\contents#1#2{\vskip 0mm \noindent{\bfseries #1} \dotfill\ #2\\}
\def\subcontents#1#2{\hbox to 1em{}{#1} \dotfill\ #2\\}
\noindent\small
\input generated/supplementary.toc
}

\newwrite\contentsfile
\immediate\openout\contentsfile=generated/supplementary.toc

@@

\vbox{\noindent
      \input generated/minicontents.tex
      }

\newwrite\minicontentsfile
\immediate\openout\minicontentsfile=generated/minicontents.tex

@@

{The type and use of mathematics used in Software Engineering is called \textbf{formal methods}. Not using formal methods ensures the resulting code is unsafe and unreliable. Of particular relevance to scientific modeling: there must be an explicit use of formal methods to ensure mathematical models (such as differential equations) are correctly implemented in code (and to understand the any limitations of doing so).}

It is important to be clear that formal methods is a spectrum, from doing it as rigorously and comprehensively as possible using state of the art methods, to applying a basic ``formal methods mindset'' that any competent programmer could do.

\begin{itemize}\raggedright\item An interesting paper explores a formal approach to coding differential equations \cite{differential-equations}, but it makes it clear that their approach is beyond most programmers. Such formal methods require sophisticated knowledge of logic \cite{cbc}, as well as practical knowledge of using appropriate formal methods tools (Alloy, HOL, PVS, SPARK \cite{spark}, and \emph{many\/} others). Using the right tools is essential for reliable programming, because the tools do quickly and reliably what, done by hand, would be slow and error-prone. Standard tools cover verification, static analysis of code, version control, documentation, and so on --- this paper explains why some of these activities are essential for reliable programming below.

\item At the other end of the spectrum, a formal methods mindset means being clear about basic mathematical properties of code, such as by using assertions, proving invariants in loops, and so on. This approach should be feasible --- and perhaps required --- for all scientific programmers.
\end{itemize}

Crucially, computer tools are available to catch common human errors that we are all prone to. Many tools are designed to avoid common human errors arising \emph{in the first place\/}; notably, the MISRA C toolset simply stops the developer using the most error-prone features of normal C, and hence improves the quality of programming with little effort \cite{misra}.

Many programming languages and programming environments have integrated features that support formal methods. For example, Hoare's triples \cite{hoare} (and formal thinking based on similar ideas) are readily supported by assertions, as either provided explicitly in a programming language or through a simple API\@. In particular, assertions readily support contracts, an important rigorous way of programming: assertions allow the program, the programming language, or tools (as the case may be) to automatically (and hence rigorously) check essential details of the program.

Hoare's original 1969 paper \cite{hoare} is very strongly recommended because it is a classic paper that has stood the test of time; in the 1960s it was leading research, but now it can be read as an excellent introduction, given how the field of Software Engineering has advanced and become more specialized and sophisticated over the decades since. Hoare is also a very good writer.

@@

\highlightsec{3}{Using inappropriate programming languages undermines reliability}{Using dependable programming languages}
Many popular languages are popular because they are easy to use, which is not the same as being reliable to use. The fewer constraints a language imposes, the easier it \emph{seems\/} to be to program in, but the lack of constraints means the language cannot provide the checks stricter languages do. C, for instance, which is one of the languages widely used for modeling \cite{tweet,plos}, is not a good choice for a reliable programming language --- it has many intrinsic weaknesses that are well-known to professionals, but which frequently trap inexperienced programmers. (This is not the place for a review of bad programming languages, for which see \cite{heedless}, but Excel is even worse than C\@.)

In particular, C is not a portable language, which means C code will work differently on different types of computer and operating system.

SPARK Ada is a popular example of a much more appropriate high integrity programming language to use \cite{spark}. SPARK Ada also has the advantage that most Ada programmers are better qualified than most C programmers.

Other high integrity languages include  OCaml, F*, and Haskell; reference \cite{haskell} is an excellent introduction to Haskell, and introduces the wider issues of reliable programming in such languages.

@@

{If results cannot be reproduced, has anything reliable been contributed? When a modeling paper presents results from a model, it is important to reproduce those results without using the same code. Better still, research should be reproduced without sharing libraries or APIs (for example, results from a model using R might be reproduced using \emph{Mathematica\/} --- this is a case of $N$ (where, in this case, $N=2$) version Programming \cite{NVP}). Reproducing the same results relying on the same codebase tells you little. The more independent reproductions of results the greater the evidence for belief in the implications.}

@@

Most version control systems would, in addition, enable open source methods so the code could be shared --- and reviewed --- by a wider community. Open source is not a panacea, however; it raises many trade-offs. Particularly for world-wide concerns like pandemic modeling, it increases diversity in the software developers, and fosters a diverse scientific collaboration. Open source can raise people's standards --- some countries \cite{excel1,excel2} are using Excel models to manage COVID-19, and as there are.are serious dependability problems with Excel (illustrated particularly in section \ref{benefits-beyond-science} of the main paper), open source projects competently implemented (e.g., avoiding or carefully managing the use of Excel) would help these people enormously.

@@

\mysection{Code, data, and publication}
\globalLabel{on-code-data-publication}

@@

This is clearly all source code, but the number \the\magicNumber\ above is likely to be some sort of relevant data, though it might be a physical constant whose value does not depend at all on \emph{this\/} experiment. The next hard-coded value mentioned in the calculation is difficult to categorize: is the value of $\pi$ empirical data or is it part of a standard formula? Some programming languages like \emph{Mathematica\/} treat $\pi$ as an exact mathematical constant (e.g., \emph{Mathematica\/} calculates $\tan\pi/4 = 1$ exactly), but $\pi$ is \emph{also\/} definitely an inexact empirical value.\footnote{A record set on 19 August 2021, the most accurate value of $\pi$ then known was 62,831,853,071,796 digits %as calculated by the University of Applied Sciences, Chur, Switzerland 
\url{www.fhgr.ch/en/specialist-areas/applied-future-technologies/davis-centre/pi-challenge}}

@@

Section \ref{over-fit} of the main paper looks like part of an ordinary paper, but it (including the figure and calculations) was data generated by a \emph{Mathematica\/} program.

@@

The idea is very simple, but very effective. The ``code'' for the paper includes the \emph{Mathematica\/} notebook that generated section \ref{over-fit}. In fact, the notebook is written as a self-contained report or paper, as \emph{Mathematica\/} notebooks generally are, and it thoroughly documents how it works.

@@

\mysubsection{Data and polynomials used in the paper}
The data and polynomials used in the main paper's section \ref{over-fit}, and illustrated in the paper's figure \ref{fig-overfit}, is presented below, as generated in \LaTeX\ by the same \emph{Mathematica\/} notebook that generated section \ref{over-fit}. Of course, in the paper itself, the data was not necessary to make the point, but if there is any need to replicate it or otherwise scrutinize the argument in the paper --- as there would be for more complex arguments in typical papers --- the data used and results are shown here. In a similar way, data and computed results could be presented and made available in other scientific papers.

\input generated/over-fitting-data.tex

\highlightmysubsection{Comparing conventional and RAP approaches}
\globalLabel{RAP-diagrams}
The similarities and differences between the conventional copy-and-paste approach to filling in data and diagrams in publications, the improved systematic RAP process, and using notebooks (such as \emph{Mathematica\/} or Jupyter), are illustrated in the sequence of schematics of figures \ref{fig:basic}, \ref{fig:excel}, \ref{fig:rap}, and \ref{fig:notebook}.

\begin{figure} 
   \centering
   \includegraphics[width=5in]{generated/basic.pdf} 
   \caption{The common basic --- error-prone and obsolete --- approach to scientific authoring is to use code to help generate analyses and diagrams, then manually copy and paste the selected results into the publication. Note that the publication, including the results, can be edited arbitrarily, and typically the results published will have been edited and modified (if only for typographical purposes) from those actually generated by the code.}
   \label{fig:basic}
\end{figure}
\begin{figure} 
   \centering
   \includegraphics[width=5in]{generated/excel.pdf} 
   \caption{When the code is in a spreadsheet, such as Microsoft Excel, the code is generally hidden from sight. The data copied \& pasted into a publication may or may not be calculated from data in other cells (such as column totals). Records are rarely taken of these manual processes, and, anyway, typically it is impossible to be certain exactly what has been copied unless very great care is taken. In consequence, if a spreadsheet is modified, it is haphazard what results are updated and corrected in the publication. }
   \globalLabelValue{excelFigure}{\thefigure}
   \label{fig:excel}
\end{figure}
\begin{figure} 
   \centering
   \includegraphics[width=5in]{generated/import.pdf} 
   \caption{Improving over the normal approach (figure \ref{fig:basic}), data and diagrams are placed in the publication using a programmed, systematic approach. In the present paper, this was done by generating text files of \LaTeX\ definitions (represented by boxes in the central column of the schematic), hence providing \LaTeX\ names for all of the code-generated values. Since the copying of results into a publication is automated and easy, any improvements to the results (such as correcting errors) are made by improving the code --- which therefore contributes to improving all future-generated results too. Not shown, but in well-engineered code, documentation will also be generated, such as by using tools such as JavaDoc or Doxygen.}
   \label{fig:rap}
\end{figure}
\begin{figure}[t]
   \centering
   \includegraphics[width=5in]{generated/notebook.pdf} 
   \caption{In a notebook system, such as \emph{Mathematica\/} or Jupyter, a single document --- the notebook --- integrates the publication, the code, any notes, as well as all the results. The code generates results and images that are automatically (and reliably) inserted in place into the notebook, usually right after the code that generates them. Some parts of the notebook are marked or selected to be printed as the publication, thus allowing most if not all code to be hidden from the final publication. In well-engineered code, the notebook will directly contain full documentation.
Note that in most systems, notebooks can import and generate arbitrary data (images, sounds, sensor data, etc).}
   \label{fig:notebook}
\end{figure}

@@

David Spiegelhalter is concerned how statistics is often misused and misunderstood. In his \emph{The Art of Statistics\/} \cite{Spiegelhalter} Spiegelhalter brings together his advice for making reliable statistical claims: they need to be accessible, intelligible, assessable, and usable --- and the claims need to be properly accountable.

Spiegelhalter proposes ten questions to ask when confronted with any claim based on statistical evidence. Some of his questions are quite general, and might be applied to any sort of scientific claims, but all have analogous questions that could be addressed to software code or publications relying on code --- analogues are suggested in \textbf{bold} below.

What might seem like dauntingly technical software issues are no more demanding than the basic statistical issues that are regularly acceded to; failing to ask these questions is as risky as dismissing statistical scrutiny.

@@

\sequestion{Cherry picking with code is often unconscious and is very common: when running code produces the ``cherries'' for a paper it is tempting to stop testing the code, and just assume it is running correctly. So, what evidence is there that the code was rigorously developed and cherry picking avoided?}

@@

The main paper was motivated by selected epidemiological papers and their problematic computational models (which are discussed in the paper). Are the issues illustrated by these case studies typical of science more broadly? We undertook, then, some selected studies of papers in different fields, then undertook a randomized, stratified pilot survey covering several leading peer reviewed journals. The point was not to establish the frequency of problems, so much as to sign whether the problematic case study was exceptional or typical.

@@

The pilot study follows the RAP methodology. The data and code for this paper's pilot survey (and all other analysis used in the paper and in this \supplement) are available on GitHub. All raw data is converted into \LaTeX\ so that the analysis can be typeset directly in the paper; this \supplement\ also contains a complete tabular presentation of the data.

Moreover, this paper itself follows the more general \RAPstar\ methodology. For example, as standard practice, a Unix makefile is used to make it easy to analyze and generate all data, typeset the paper, and more. Table \ref{table:options} shows the options provided.

\begin{table}[t]
\begin{center}
\input generated/make-help.tex
\end{center}
\caption{Conforming to the \RAPstar\ methodology, the abbreviated summary above was generated automatically, by using \texttt{make data}. %(Saying \texttt{make help} would generate a full summary on the command line.)
}
\label{table:options}
\end{table}

The \RAPstar\ approach cannot enforce the truth of such as summary, but it very significantly increases the chances that the summary is correct and up to date. For example, if refactoring leads to an option being deleted, then it will also disappear \emph{with no further work\/} from the table above. Also, since the summary was proof-read at the same time as proof-reading this \supplement, following \RAPstar\ also increases the chances that any errors or functionality omissions or issues in the makefile have been detected and corrected.

@@

%\newcount \total
%\total=\dataN
%\multiply \total by 4 % three reviewers, one editor
%\advance \total by \countAuthors
%\the\total\
%leading scientists assuming authors, 1 editor, and 3 reviewers per paper, but not counting acknowledged colleagues et al, covering  (plus appendices and supporting material) in total.
%
\caption{Overview of the peer-reviewed paper sample.}
\label{supplement-table-overview}
\end{table*}

\begin{table*}[t]
\begin{center}\normalsize
\input generated/summary-table.tex
\end{center}

@@

% divide #1 by #2
\newcount \uu
\newcount \vv
\newcount \ww
\def\makeAverage#1#2{%
	\uu=#1
	\vv=#2
	\multiply \uu by 1000
	\divide \uu by \vv
	% uu is now 1000 times the answer we want
	\ww = \uu
	\divide \uu by 1000
	\multiply \uu by 1000
	\advance \ww by -\uu
	\divide \uu by 1000
	%[ #1 / #2 = \the\uu .\the\ww\ ]
	\ifnum \ww > 499 \advance \uu by 1 \fi
	\the\uu
}

@@

\input generated/repos.tex
\vskip 1ex
\small Citation numbers $>$ \MaxMainPaperCitationNumber\ can be found in the \supplement
\\
\small Repository clones downloaded and automatically summarized \clonedate
\end{center}

\caption{Sizes of repositories, with approximate sizes of code (in kLOC) and data for all available GitHub repositories reviewed in the survey, plus \texttt{covid-sim} \cite{ICmodel} for comparison. Sizes are approximate because in all repositories code and data are conceptually interchangeable (an issue explained in the \supplement), so choices were made in the survey to avoid double-counting. Many repositories rely on downloading additional code and data, which is not counted in the table as the additional required material is not in the repository cited in the paper.  At the time of cloning and checking all repositories in \clonewhen, paper \csname cite-PostoperativeOutcomes.RiskNet\endcsname\ still had nothing in its repository except a single file still saying ``...code coming soon...,'' despite \the\pubdelayinmonths\ months having already elapsed since the submitted paper had claimed the code could be accessed in its repository.}
\label{table-repo-summary}\globalTableLabel{supplementaryTableRepoSummary}
\end{table*}

@@

With the one exception, none of the papers used any specific Software Engineering methods, such as open source \cite{open-source} or other standard methodologies provided in this \supplement, to help manage their processes and help improve quality. Although software stability \cite{stability} is a relatively new concept, understood as methodologies, such as portability, to provide long-term value of software, it is curious that none of the papers made any attempt at stability (however understood) despite the irony that all the papers were published in archival journals.\footnote{Reasons the present paper does not directly assess the quality of software in the surveyed papers include: many papers did not provide complete software; it was not possible to find correct versions of all software systems to run the models; also, no papers provided adequate test suites so that correct operation of software could be confirmed objectively.}

@@

Overall, there was no evidence that any code had been developed carefully, let alone by using recognized professional Software Engineering methods. In particular, 
\ifnum \countCodetested=0
no papers
\else only \plural{\countCodetested}{paper}
\fi in the survey provide any claims or evidence of effective testing, for instance with evidence that tests were run on clean builds. {While it may sound unrealistic to ask for evidence on software quality in a paper written for another field of science, the need is no less than the need for standard levels of rigor in statistics reporting, as discussed in the opening of this paper.}

@@

It is noteworthy that none of the journals sampled permit any reliable style of managing data in published papers, such as described above in sections \ref{detecting-against-error} and \ref{restrictive-policies}. In particular, for all the papers that had accessible code, the code included explicit (and relevant) data that was not archived \emph{as\/} data in the journal repositories.

Note that journal policies were first accessed on 29 July 2020, close after the period covered by the pilot survey, thus presumably fairly closely reflecting the policies in use for the papers in the survey on the dates when they were each submitted. Unfortunately for the survey, the journals do not make clear what exact policies were applied to each paper when the papers were submitted (on the other hand, the survey shows that policies are not rigorously enforced).
\\

@@

\specifyurl
\url{royalsociety.org/journals/authors/author-guidelines/#data}
\specifyurletc{Since first accessed 29 July 2020, the policy has been revised (undated, accessed 2 February 2022) but retains the same principles; full policy now available via a DOI \cite{RS-policy}. The policy still retains an emphasis on data accessibility, and continues a lack of awareness that code and data are equivalent and often mixed (see section \ref{on-code-data-publication}).}

@@

\specifyurl
\url{www.nature.com/nature-research/editorial-policies/reporting-standards#availability-of-data}
\specifyurletc{Accessed 29 July 2020; since updated (accessed 2 February 2022) to require [in part] ``Upon publication, Nature Portfolio journals consider it best practice to release custom computer code in a way that allows readers to repeat the published results. Code should be deposited in a DOI-minting repository such as Zenodo, Gigantum or Code Ocean and cited in the reference list following the guidelines described here.'' 
}

@@

\specifyurl
\url{marlin-prod.literatumonline.com/pb-assets/Lancet/authors/tldh-info-for-authors.pdf}
\specifyurletc{Accessed 29 July 2020. Still no code policy when accessed 2 February 2022.}

\noindent\textbf{Extract from \emph{Journal of Vascular Surgery\/} author guidelines}\\
The \emph{Journal of Vascular Surgery\/} has detailed data policies, but no code policy. While no \emph{Journal of Vascular Surgery\/} papers were surveyed, the following statement on data policies is relevant:

@@

\newdimen \codeIndent
\codeIndent = 1em
\vskip 1ex
\vbox{%
  \input generated/example-data.tex
}
\vskip 1ex

@@

The JavaScript data+program generates files from the JSON; these files were then included in both the main paper and in this \supplement, so when the paper or \supplement\ is typeset all tables and specific data items are typeset automatically, consistently and reliably by \LaTeX\@.

For example, the register \texttt{\bslash dataN} is set to the value \the\dataN, which is the total number of papers assessed in the JSON data, and the macro \texttt{\bslash journalBreakdown} is defined directly from the data to be the following text (when typeset in \LaTeX):

@@

\noindent
\vbox{\noindent Legend:\vskip 1ex
\input generated/legend.tex
}
%\newpage

>>>>

\iftrue
    \PassOptionsToPackage{hyphens}{url}
    \documentclass[10pt,a4paper]{article}
    \usepackage[text={6.5in,10in},centering]{geometry}    
    \usepackage{url}
    \usepackage{doi}
    \usepackage{longtable}
    \usepackage{color}
    \usepackage{amsmath,amssymb}
    \usepackage{fancyhdr}
    \def\httpURL#1{\href{http://#1}{\textcolor{blue}{#1}}}
    \def\LONGhttpURL#1#2{\href{http://#1}{\textcolor{blue}{#2}}}    
    \input paper-seb-macros.tex
\fi

@@

\input generated-page-lengths.tex

@@

% pageref gives us a hyperlink which looks unsightly
\makeatletter
\def\rememberpage{\immediate\write\@auxout{\bslash gdef\bslash SupplementalLastPage{\thepage}}}
\makeatother

@@

% this will be used to save cross references that can be used in the main paper too
\newwrite\infofileForMain
\immediate\openout\infofileForMain=generated-info-for-main.tex

\input generated-info-for-supplement.tex
\newwrite\infofileForSupplement
\immediate\openout\infofileForSupplement=generated-info-for-supplement.tex

@@

\makeatletter
\def\globalLabel#1{% we also need journal code-policies to be used in the main paper...
\typeout{Global xref: #1}%
\immediate\write\infofileForMain{\bslash newlabel{#1}{{\arabic{section}}{\thepage}{DUMMYA}{DUMMYB}{DUMMYC}}}
% as well as in the current document....
\immediate\write\@auxout{\bslash newlabel{#1}{{\bslash arabic{section}\dotalph{subsection}}{\thepage}{AAA}{BBB}{CCC}}}
}
\makeatother

@@

\noindent\vbox{% The table of contents
\def\contents#1#2{\vskip 0mm \noindent{\bfseries #1} \dotfill\ #2\\}
\def\subcontents#1#2{\hbox to 1em{}{#1} \dotfill\ #2\\}
\noindent\small
\input generated-supplementary.toc
}

\newwrite\contentsfile
\immediate\openout\contentsfile=generated-supplementary.toc

@@

\vbox{\noindent
      \input generated-minicontents.tex
      }

\newwrite\minicontentsfile
\immediate\openout\minicontentsfile=generated-minicontents.tex

@@

{The type and use of mathematics used in Software Engineering is formal methods. Not using formal methods ensures the resulting code is unsafe and unreliable. Of particular relevance to scientific modeling: there must be an explicit use of formal methods to ensure  mathematical models (such as differential equations) are correctly implemented in code (and to understand the any limitations of doing so).}

Formal methods require sophisticated knowledge of logic \cite{cbc}, as well as practical knowledge of using appropriate formal methods tools (Alloy, HOL, PVS, SPARK, and \emph{many\/} others). Using the right tools is essential for reliable programming, because the tools do quickly and reliably what, done by hand, would be slow and error-prone. Standard tools cover verification, static analysis of code, version control, documentation, and so on --- this paper explains why some of these activities are essential for reliable programming below.

Crucially, tools are designed to catch common human errors that we are all prone to. Many tools are designed to avoid common human errors arising \emph{in the first place\/}; notably, the MISRA C toolset simply stops the developer using the most error-prone features of normal C, and hence improves the quality of programming with little effort \cite{misra}.

Many programming languages and programming environments have integrated features that support formal methods. For example, Hoare's triples \cite{hoare} (and formal thinking based on similar ideas) are readily supported by assertions, as either provided explicitly in a programming language or through a simple API\@. In particular, assertions readily support contracts, an important rigorous way of programming: assertions allow the program, the programming language, or tools (as the case may be) to automatically (and hence rigorously) check essential details of the program. Hoare's original 1969 paper \cite{hoare} is very strongly recommended because it is a classic paper that has stood the test of time; in the 1960s it was leading research, but now it can be read as an excellent introduction, given how the field of Software Engineering has advanced and become more specialized and sophisticated over the decades since. Hoare is also a very good writer.

@@

\highlightsec{3}{Using inappropriate programming languages undermines reliability}{Using dependable programming languages}
Many popular languages are popular because they are easy to use, which is not the same as being reliable to use. The fewer constraints a language imposes, the easier it \emph{seems\/} to be to program in, but the lack of constraints means the language cannot provide the checks stricter languages do. C, for instance, which is one of the languages widely used for modeling \cite{tweet,plos}, is not a good choice for a reliable programming language --- it has many intrinsic weaknesses that are well-known to professionals, but which frequently trap inexperienced programmers. (This is not the place for a review \cite{heedless} but Excel is even worse in this regard.) In particular, C is not a portable language (unless extreme care is taken), which means models will work differently on different types of computer. SPARK Ada is one example of a much more appropriate programming language to use. SPARK Ada also has the advantage that most Ada programmers are better qualified than most C programmers.

@@

{If results cannot be reproduced, has anything reliable been contributed? When a modeling paper presents results from a model, it is important to reproduce those results without using the same code. Better still, research should be reproduced without sharing libraries or APIs (for example, results from a model using R might be reproduced using Mathematica --- this is a case of $N$ (where, in this case, $N=2$) version Programming \cite{NVP}). Reproducing the same results relying on the same codebase tells you little. The more independent reproductions of results the greater the evidence for belief in the implications.}

@@

Most version control systems would, in addition, enable open source methods so the code could be shared --- and reviewed --- by a wider community. Open source is not a panacea, however; it raises many trade-offs. Particularly for world-wide concerns like pandemic modeling, it increases diversity in the software developers, and fosters a diverse scientific collaboration. Open source can raise people's standards --- some countries \cite{excel1,excel2} are using Excel models to manage COVID-19, and open source projects properly implemented would help these people enormously.

@@

\mysection{Code, data and publication}
\globalLabel{on-code-data-publication}

@@

This is clearly all source code, but the number \the\magicNumber\ above is likely to be some sort of relevant data, though it might be a physical constant whose value does not depend at all on \emph{this\/} experiment. The next hard-coded value mentioned in the calculation is difficult to categorize: is the value of $\pi$ empirical data or is it part of a standard formula? Some programming languages like Mathematica treat $\pi$ as an exact mathematical constant (e.g., Mathematica calculates $\tan\pi/4 = 1$ exactly), but $\pi$ is \emph{also\/} definitely an inexact empirical value.\footnote{A record set on 19 August 2021, the most accurate value of $\pi$ then known was 62,831,853,071,796 digits %as calculated by the University of Applied Sciences, Chur, Switzerland 
\url{www.fhgr.ch/en/specialist-areas/applied-future-technologies/davis-centre/pi-challenge}}

@@

Section \ref{over-fit} of the main paper looks like part of an ordinary paper but in fact it was generated by a \emph{Mathematica\/} program. The section looks like typical text from a paper, but it was data generated by a program.

@@

The idea is very simple, but very effective. The ``code'' for this paper includes the \emph{Mathematica\/} notebook that generated section \ref{over-fit}. In fact, the notebook is written as a self-contained report or paper, as \emph{Mathematica\/} notebooks generally are, and it thoroughly documents how it works.

@@

@@

David Spiegelhalter is concerned how statistics is often misused and misunderstood. In his \emph{The Art of Statistics\/} \cite{Spiegelhalter} Spiegelhalter brings together his advice for making reliable statistical claims: they need to be accessible, intelligible, assessable, and usable --- the claims need to be properly accountable. Spiegelhalter proposes ten questions to ask when confronted with any claim based on statistical evidence. Some of his questions are quite general, and might be applied to any sort of scientific claims, but all have analogous questions that could be addressed to software code or papers relying on code --- analogues are suggested in \textbf{bold} below.

What might seem like dauntingly technical software issues are no more demanding than the basic statistical issues that are regularly acceded to; failing to ask them is as risky as dismissing statistical scrutiny.

@@

\sequestion{Cherry picking with code is often unconscious and is very common: when running code produces the ``cherries'' for a paper it is tempting to stop testing the code and just assume it is running correctly. So, what evidence is there that the code was rigorously developed and cherry picking avoided?}

@@

The main paper focuses on selected epidemiological papers and their computational models. Are the issues illustrated by these case studies typical of computational science more broadly? We undertook, first, some selected studies of papers in different fields, then undertook a randomized, stratified pilot survey covering several leading peer reviewed journals.

@@

The pilot study follows the RAP methodology. The data and code for this paper's pilot survey (and all other analysis used in the paper and in this \supplement) are available on Github. All raw data is converted into \LaTeX\ so that the analysis can be typeset directly in the paper; this \supplement\ also contains a complete tabular presentation of the data.

Moreover, this paper follows the more general \RAPstar\ methodology. For example, as standard practice, a Unix makefile is used to make it easy to analyze and generate all data, typeset the paper, and more. The following options are provided:

\vskip 1ex{\small \begin{center}
\input generated.make-help.tex
\end{center}
}\vskip 1ex

Conforming to the \RAPstar\ methodology, the summary above (including all the \LaTeX\ typesetting details) was generated automatically (in fact, by just saying ``\texttt{make data}'', which also generated all the other data). However, when \texttt{make help} or \texttt{make help-brief} are run, they generate plain text summaries without any of the \LaTeX\ typesetting details used here (e.g., \LaTeX\ will appear simply as \texttt{Latex}).

The \RAPstar\ approach cannot enforce the truth of the summary, but it very significantly increases the chances that the option summary is correct and up to date. For example, if refactoring leads to an option being deleted, then it will also disappear \emph{with no further work\/} from the table above. Also, since the summary was proof-read at the same time as proof-reading this \supplement, following \RAPstar\ also increases the chances that any errors or functionality omissions or issues in the makefile have been detected and corrected.

@@

%\newcount \total
%\total=\dataN
%\multiply \total by 4 % three reviewers, one editor
%\advance \total by \countAuthors
%\the\total\
%leading scientists assuming authors, 1 editor, and 3 reviewers per paper, but not counting acknowledged colleagues et al, covering  (plus appendices and supporting material) in total.
%
\caption{Overview of peer-reviewed paper sample.}
\label{supplement-table-overview}
\end{table*}

\begin{table*}[t]
\begin{center}\normalsize
\input generated-summary-table.tex
\end{center}

@@

% divide #1 by #2
\newcount \uu
\newcount \vv
\def\makeAverage#1#2{%
	\uu=#1
	\vv=#2
	\divide \uu by \vv
	\the\uu
}

@@

\input generated-repos.tex
\vskip 1ex
\small Citation numbers $>$ \MaxMainPaperCitationNumber\ can be found in the \supplement
\\
\small Repository clones downloaded and automatically summarized \clonedate
\end{center}

\caption{Sizes of repositories, with approximate sizes of code (in kLOC) and data for all available GitHub repositories reviewed in the survey, plus \texttt{covid-sim} \cite{ICmodel} for comparison. Sizes are approximate because in all repositories code and data are conceptually interchangeable (an issue explained in the \supplement), so choices were made in the survey to avoid double-counting. Many repositories rely on downloading additional code and data, which is not counted in the table as the additional required material is not in the repository cited in the paper.  At the time of cloning and checking all repositories in \clonewhen, paper \csname cite-PostoperativeOutcomes.RiskNet\endcsname\ still had nothing in its repository except a single file still saying ``...code coming soon...,'' despite \the\pubdelayinmonths\ months having already elapsed since the submitted paper had claimed the code could be accessed in its repository.}
\label{table-repo-summary}
\end{table*}

@@

With the one exception, none of the papers used any specific Software Engineering methods, such as open source \cite{open-source} or other standard methodologies provided in the \supplement, to help manage their processes and help improve quality. Although software stability \cite{stability} is a relatively new concept, understood as methodologies, such as portability, to provide long-term value of software, it is curious that none of the papers made any attempt at stability (however understood) despite the irony that all the papers were published in archival journals.\footnote{Reasons the present paper does not directly assess the quality of software in the surveyed papers include: many papers did not provide complete software; it was not possible to find correct versions of all software systems to run the models; also, no papers provided adequate test suites so that correct operation of software could be confirmed objectively.}

@@

Overall, there was no evidence that any code had been developed carefully, {let alone by using recognized professional Software Engineering methods}. In particular, 
\ifnum \countCodetested=0
no papers
\else only \plural{\countCodetested}{paper}
\fi in the survey provide any claims or evidence of effective testing, for instance with evidence that tests were run on clean builds. {While it may sound unrealistic to ask for evidence on software quality in a paper written for another field of science, the need is no less than the need for standard levels of rigor in statistics reporting, as discussed in the opening of this paper.}

@@

It is noteworthy that none of the journals sampled permit any reliable style of managing data in published papers, such as described above in sections \ref{detecting-against-error} and \ref{restrictive-policies}. In particular, for all the papers that had accessible code, the code included explicit (and relevant) data that was not archived \emph{as\/} data in the journal repositories.
\\

@@

\specifyurl
\url{royalsociety.org/journals/authors/author-guidelines/#data}
\specifyurletc{Accessed 29 July 2020; the policy has been revised (undated, but accessed 2 February 2022) but retains the same principles; full policy now available via a DOI \cite{RS-policy}. The policy still retains an emphasis on data accessibility, and continues a lack of awareness that code and data are equivalent and often mixed (see section \ref{on-code-data-publication}).}

@@

\specifyurl
\url{www.nature.com/nature-research/editorial-policies/reporting-standards#availability-of-data}
\specifyurletc{Accessed 29 July 2020; since updated to require [in part] ``Upon publication, Nature Portfolio journals consider it best practice to release custom computer code in a way that allows readers to repeat the published results. Code should be deposited in a DOI-minting repository such as Zenodo, Gigantum or Code Ocean and cited in the reference list following the guidelines described here.'' (accessed 2 February 2022).
}

@@

\specifyurl
\url{marlin-prod.literatumonline.com/pb-assets/Lancet/authors/tldh-info-for-authors.pdf}
\specifyurletc{Accessed 29 July 2020. Still no code policy (accessed 2 February 2022).}

\noindent\textbf{Extract from \emph{Journal of Vascular Surgery\/} author guidelines}\\
Journal has detailed data policies, but no code policy. While no \emph{Journal of Vascular Surgery\/} papers were surveyed, the following statement on data policies is relevant:

@@

\newdimen \codeIndent
\codeIndent = 1em
\vskip 1ex
\vbox{%
  \input generated-example-data.tex
}
\vskip 1ex

@@

The JavaScript data+program generates files from the JSON, with all the definitions; these files were then included in both the main paper and in this \supplement, so when the paper or \supplement\ is typeset all tables and specific data items are typeset automatically, consistently and reliably by \LaTeX\@.\footnote{Because of its approach to automatic typesetting, this journal requires that no files are explicitly \texttt{input} into \LaTeX, so a simple JavaScript program is used to recursively expand all \texttt{input} commands before submission. If the source files are available for download after this paper is accepted, they will therefore contain no \texttt{input} commands, but they will contain comments at the appropriate points in the expanded source code explaining the sources of the \texttt{input} data.} For example, the register \texttt{\bslash dataN} is set to the value \the\dataN, which is the total number of papers assessed in the JSON data, and the macro \texttt{\bslash journalBreakdown} is defined directly from the data to be the following text (when typeset in \LaTeX):

@@

\noindent
\vbox{\noindent Legend:\vskip 1ex
\input generated-legend.tex
}
%\newpage
